---
title: "boxplots"
author: "Eunice Yeh"
date: "April 21, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Import RMP dataset and subset
- use derived dataset of RMP abundances for the two participants of interest (different health states)
- subset to the two specific bacteria of interest: Prevotella copri and Escherichia coli (E.coli) -- both at the genus and species level
- create a column to group/distinguish the two participants (health states)

```{r import-subset}
RMP <- read.table(file = "../data/derived/transposed_data_subset_genus_or_species_level.tsv", sep = '\t', header = TRUE) %>% 
  select(X, Prevotella = ends_with("g__Prevotella"), Escherichia = ends_with("g__Escherichia"),     # genus level
            P_copri = ends_with("s__Prevotella_copri"), E_coli = ends_with("s__Escherichia_coli")) %>% # species level
  mutate(Diagnosis = c(rep("H4023 (Control)",22),rep("H4015 (CD)",22))) # the first 22 samples are for H4023 (control), rest are for H4015 (CD)
head(RMP, 44)
```

## Import QMP dataset and subset
- use derived dataset of QMP abundances (that has been calculated in a separate program)
- subset to the same two specific bacteria of interest
- again, create the grouping variable for the diagnosis

```{r import-QMP}
QMP_CD <- read.csv("../data/derived/qmp4015final.csv", header=TRUE) %>% 
  select(X, P_copri = ends_with("s__Prevotella_copri"), E_coli = ends_with("s__Escherichia_coli")) %>% 
  mutate(Diagnosis = rep("H4015 (CD)"))

QMP_con <- read.csv("../data/derived/qmp4023final.csv", header=TRUE) %>% 
  select(X, P_copri = ends_with("s__Prevotella_copri"), E_coli = ends_with("s__Escherichia_coli")) %>% 
  mutate(Diagnosis = rep("H4023 (Control)"))

QMP <- bind_rows(QMP_CD, QMP_con) # combine the two QMP datasets into one
```


## Boxplot by health state for each bacteria in RMP vs. QMP

- first in RMP

```{r RMP-pcopri-boxplots}
# prevotella RMP
# genus level
RMP %>% ggplot(aes(Diagnosis, Prevotella, fill=Diagnosis), group=Diagnosis) + 
  geom_boxplot() + geom_jitter(width = 0.02) + 
  scale_y_continuous(trans = "log10") + 
  ylab("Relative Genus Abundance (RMP)") +
  theme_bw() + theme(legend.position="none") +
  ggtitle("Prevotella")
# species level
RMP %>% ggplot(aes(Diagnosis, P_copri, fill=Diagnosis), group=Diagnosis) + 
  geom_boxplot() + geom_jitter(width = 0.02) + 
  scale_y_continuous(trans = "log10") + 
  ylab("Relative Species Abundance (RMP)") +
  theme_bw() + theme(legend.position="none") +
  ggtitle("Prevotella copri")
```

```{r RMP-ecoli-boxplots}
# e. coli RMP
# genus level
RMP %>% ggplot(aes(Diagnosis, Escherichia, fill=Diagnosis), group=Diagnosis) + 
  geom_boxplot() + geom_jitter(width = 0.02) + 
  scale_y_continuous(trans = "log10") + 
  ylab("Relative Genus Abundance (RMP)") +
  theme_bw() + theme(legend.position="none") +
  ggtitle("Escherichia")
# species level
RMP %>% ggplot(aes(Diagnosis, E_coli, fill=Diagnosis), group=Diagnosis) + 
  geom_boxplot() + geom_jitter(width = 0.02) + 
  scale_y_continuous(trans = "log10") + 
  ylab("Relative Species Abundance (RMP)") +
  theme_bw() + theme(legend.position="none") +
  ggtitle("Escherichia coli")
```

- then in QMP (only focusing on species level)

```{r QMP-boxplots}
# prevotella copri QMP
QMP %>% ggplot(aes(Diagnosis, P_copri, fill=Diagnosis), group=Diagnosis) + 
  geom_boxplot() + geom_jitter(width = 0.02) + 
  scale_y_continuous(trans = "log10") + 
  ylab("Quantitative Species Abundance (QMP, ng/uL)") +
  theme_bw() + theme(legend.position="none") +
  ggtitle("Prevotella copri")

# e. coli QMP
QMP %>% ggplot(aes(Diagnosis, E_coli, fill=Diagnosis), group=Diagnosis) + 
  geom_boxplot() + geom_jitter(width = 0.02) + 
  scale_y_continuous(trans = "log10") + 
  ylab("Quantitative Species Abundance (QMP, ng/uL)") +
  theme_bw() + theme(legend.position="none") +
  ggtitle("Escherichia coli")

```

- Saving images 500 x 500 for presentation

