---
title: "Log10-transformed t tests"
author: "Eunice Yeh, Anthony Lamattina"
date: "April 21, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Import RMP dataset and subset
- use derived dataset of RMP abundances for the two participants of interest (different health states)
- subset to the two specific bacteria of interest: Prevotella copri and Escherichia coli (E.coli) -- both at the genus and species level
- create a column to group/distinguish the two participants (health states)
- log10 transform the values

```{r import-subset}
RMP <- read.table(file = "data/derived/transposed_data_subset_genus_or_species_level.tsv", sep = '\t', header = TRUE) %>% 
  select(X, Prevotella = ends_with("g__Prevotella"), Escherichia = ends_with("g__Escherichia"),     # genus level
            P_copri = ends_with("s__Prevotella_copri"), E_coli = ends_with("s__Escherichia_coli")) %>% # species level
  mutate(Diagnosis = c(rep("H4023 (Control)",22),rep("H4015 (CD)",22))) # the first 22 samples are for H4023 (control), rest are for H4015 (CD)
head(RMP, 44)
```

## Import QMP dataset and subset
- use derived dataset of QMP abundances (that has been calculated in a separate program)
- subset to the same two specific bacteria of interest
- again, create the grouping variable for the diagnosis

```{r}

QMP_CDfinal <- read.csv("data/derived/qmp4015final.csv", header=TRUE) %>% select(X, P_copri = ends_with("s__Prevotella_copri"), E_coli = ends_with("s__Escherichia_coli")) %>% mutate(Diagnosis = rep("H4015 (CD)"))

QMP_Confinal <- read.csv("data/derived/qmp4023final.csv", header=TRUE) %>% select(X, P_copri = ends_with("s__Prevotella_copri"), E_coli = ends_with("s__Escherichia_coli")) %>% mutate(Diagnosis = rep("H4023 (Control)"))

QMPfinal <- bind_rows(QMP_CDfinal, QMP_Confinal) # combine the two QMP datasets into one

head(QMPfinal, 44)
```


## Log10-transform data and T test

- first in RMP

```{r log10RMP}
# log10 transform RMP data
RMPlog <- RMP
RMPlog[,2:5] <- log10(RMPlog[,2:5])

# subset logRMP into CD and Control dataframes
RMPlogCD <- subset(RMPlog, Diagnosis == "H4015 (CD)")
RMPlogCon <- subset(RMPlog, Diagnosis == "H4023 (Control)")

# transformed t tests (Welch Two Sample t-test), ignoring infinite values

# P. copri
t.test(RMPlogCD$P_copri[!is.infinite(RMPlogCD$P_copri)], RMPlogCon$P_copri[!is.infinite(RMPlogCon$P_copri)])
# P = 0.02426

# E. coli
t.test(RMPlogCD$E_coli[!is.infinite(RMPlogCD$E_coli)], RMPlogCon$E_coli[!is.infinite(RMPlogCon$E_coli)])
# P = 0.0001723
```

- now in QMP (copy number corrected)

```{r log10QMP}

# for copy-number corrected data #

# log10 transform QMP data
QMPlog <- QMPfinal
QMPlog[,2:3] <- log10(QMPlog[,2:3])

# subset logRMP into CD and Control dataframes
QMPlogCD <- subset(QMPlog, Diagnosis == "H4015 (CD)")
QMPlogCon <- subset(QMPlog, Diagnosis == "H4023 (Control)")

# transformed t tests (Welch Two Sample t-test), ignoring infinite values

# P. copri
t.test(QMPlogCD$P_copri[!is.infinite(QMPlogCD$P_copri)], QMPlogCon$P_copri[!is.infinite(QMPlogCon$P_copri)])
# P = 0.06019

# E. coli
t.test(QMPlogCD$E_coli[!is.infinite(QMPlogCD$E_coli)], QMPlogCon$E_coli[!is.infinite(QMPlogCon$E_coli)])
# P = 0.004791
```

